Simulation connectance richness

### 2023-05-14 report/cs_analysis.Rmd
Add targets for result analysis

```{r, results = FALSE}
library(tidyverse)
library(magrittr)
library(cowplot)
library(here)
library(arrow)
library(piecewiseSEM)
library(semEff)
library(easystats)
library(glmmTMB)

@@

```{r}
sim_files <- list.files(here("res", "simCSZenrich"))
sim_list <- map(sim_files, function (x) {
  open_dataset(here::here("res", "simCSZenrich", x),
    format = "arrow") %>%
  collect()
})
```

```{r}
# TODO set extinction threshold in simulations
simZ_prep <- map_dfr(sim_list,
  function (x) {
    x %>%
      # Remove simulation that have failed to run
      filter(!is.na(richness)) %>%
      # Transform back interaction strength into matrices
      mutate(across(where(is.list), as.list)) %>%
      mutate(across(c(int_strength, max_int),
          ~map2(.x, S,
            function(y, rich) {
              matrix(y, nrow = rich)
            }
          )
      )
        ) %>%
      # Get matrix of alive species
      mutate(max_int_alive  = map2(max_int, bm_sp,
          function(mat, bm) {
            mask <- bm > 10^-6
            mat[mask, mask, drop = FALSE]
          }
        )
        )
  }
)
# Compute metrics of interest
simZ <- simZ_prep %>%
  mutate(
    persistence = richness / S,
    async = 1 / sync,
    stab_pop = 1 / avg_cv_sp,
    ct_alive = map_dbl(max_int_alive,
      ~ifelse(sum(.x) == 0, 0,
      sum(.x > 0) / ((ncol(.x))^2))),
    bm_sp_alive = map(bm_sp, ~.x[.x > 10^-6]),
    tlvl_alive = map2(tlvl, bm_sp, ~.x[.y > 10^-6]),
    max_tlvl_alive = map_dbl(tlvl_alive, max),
    w_avg_tlvl_alive = map2_dbl(tlvl_alive, bm_sp_alive, ~sum(.x * (.y / sum(.y)))),
    max_max_int_alive = map_dbl(max_int_alive,
      ~ifelse(length(.x[.x > 0]) == 0, 0, max(.x[.x > 0]))
      ),
    avg_max_int_alive = map_dbl(max_int_alive,
      ~ifelse(length(.x[.x > 0]) == 0, 0, mean(.x[.x > 0]))
      ),
    gini_max_int_alive = map_dbl(max_int_alive, ~gini(.x))
      )
simZ$tlvl_alive[[3]]
simZ$max_int_alive[[3]]
```

@@

@@

```{r}
p_basic <- ~corrplot::corrplot(
  cor(simZ %>%
    filter(productivity == "enrichment") %>%
    select(c(S, richness, ct, Z, ct_alive, total_biomass, avg_max_int_alive,
        w_avg_tlvl_alive, max_tlvl_alive, rho, env_stoch, stab_com, avg_cv_sp,
        sync)),
    method = "spearman"),
  diag = FALSE, type = "upper"
)
p_enrich <- ~corrplot::corrplot(
  cor(simZ %>%
    filter(productivity == "basic") %>%
    select(c(S, richness, ct, Z, ct_alive, total_biomass, avg_max_int_alive,
        w_avg_tlvl_alive, max_tlvl_alive, rho, env_stoch, stab_com, avg_cv_sp,
        sync)),
    method = "spearman"),
  diag = FALSE, type = "upper"
)
plot_grid(
  ggdraw(p_basic),
  ggdraw(p_enrich),
  nrow = 2
)
```

@@

@@

```{r}
sim_rho <- simZ %>%
  #filter(ct_alive != 0) %>%
  mutate(across(c(async, sync, stab_pop, avg_cv_sp, stab_com, richness, avg_max_int_alive, gini_max_int), log, , .names = "log_{.col}")) %>%
  mutate(
    log1_avg_max_int_alive = log(avg_max_int_alive + 1),
    productivity_f = factor(productivity)

    ) %>%
  select(c(
      ct_alive, log_richness, max_tlvl_alive, Z, rho,
      gini_max_int, avg_max_int_alive, log1_avg_max_int_alive, env_stoch,
      log_stab_com, log_sync, log_async, log_stab_pop, log_avg_cv_sp, productivity, productivity_f,
      persistence)
  )
sim_rho_scaled <- sim_rho %>%
  mutate(across(where(is.double), ~scale(.x)[,1]))
```

@@

>>>>

```{r, results = FALSE}
library(tidyverse)
library(magrittr)
library(ggcorrplot)
library(cowplot)
library(here)
library(arrow)
library(piecewiseSEM)
library(semEff)
library(easystats)
library(glmmTMB)
library(targets)

library(rpart)
library(rpart.plot)
library(caret)
library(ipred)

@@

@@

# Introduction

- Response diversity is key in driving stability in ecological communities
- Response diversity control asynchrony with richness
- Mainly done in plants
- Food-webs structure can also modulate synchrony by predator-prey relationships
- Interaction strength, connectance, richness and enrichment can also drive more

# Research questions

##  What are the mecanisms driving stability in food-webs ?

- How structure of food-webs (connectance, species richness, prey-predator
     size ratio) gives emergence to food-web structure (max trophic level,
     weighted avg trophic level, omnivory) ?
- How food-web structure influence biomass fluxes (avg interaction
     strength, sd of interaction strength, asymetry)?
- How food-web structure and biomass fluxes affect stability components?
     total biomass, species variability, statistical averaging, evenness,
     compensatory dynamics (interaction and environment)?

# What is the effects of species richness on stability facing multiple stressors in competitive and food-webs communities?

- How do species richness drive average population stability and asynchrony
  in competitive and food-webs communities ?
- How is it affected by enrichment and response diversity?

# Material & Methods

## Bioenergetic model

TO Complete

- Important bits:
  - Niche model (Connectance highly constrained by richness)
  - Type III functional response
  - No predator interference
  - Standardize carrying capacity? (Remove the effect of species richness on
    total biomass, i.e. remove the effect of niche complementarity on biomass
    produced by primary producers)

## Variables

- S: 5 to 60
- C: .02 to .3
- K: 1 and 20
- $\rho$: 0 to 1
- Z: 10 to 1000

```{r}
tar_load(simZ)
tar_load(sim_rho)
tar_load(sim_rho_scaled)
```

@@

```{r}
cor_enrich <- cor(simZ %>%
    filter(productivity == "enrichment") %>%
    select(c(S, richness, ct, Z, ct_alive, total_biomass, avg_max_int_alive,
        w_avg_tlvl_alive, max_tlvl_alive, rho, env_stoch, stab_com, avg_cv_sp,
        sync)),
    method = "spearman")
cor_basic <- cor(simZ %>%
    filter(productivity == "basic") %>%
    select(c(S, richness, ct, Z, ct_alive, total_biomass, avg_max_int_alive,
        w_avg_tlvl_alive, max_tlvl_alive, rho, env_stoch, stab_com, avg_cv_sp,
        sync)),
    method = "spearman")
p_cor_basic <- ggcorrplot(cor_basic, type = "lower",
   lab = TRUE)
p_cor_enrich <- ggcorrplot(cor_enrich, type = "lower",
   lab = TRUE)
p_cor <- plot_grid(
  p_cor_basic,
  p_cor_enrich,
  ncol = 2
)
ggsave(here("report", "figures", "p_cor.pdf"),
  plot = p_cor,
  device = NULL,
  path = NULL,
  scale = 3.0,
  units = "mm",
  width = 140,
  height = 140 * .5,
  dpi = 300,
  limitsize = TRUE,
  bg = NULL
)
```

@@

```{r}
p_sim_rich_prod <- sim_rich_prod %>%
  filter(ct_alive != 0) %>%
  mutate(stab_pop = 1/ avg_cv_sp) %>%
  pivot_longer(
    c(stab_com, stab_pop, sync),
    names_to = "metric",
    values_to = "stab") %>%
  mutate(
    metric = var_replacement()[metric],
    rho = rho_replacement()[as.character(rho)]
  ) %>%
  ggplot(aes(x = richness, y = stab, color = factor(productivity))) +
  geom_point(alpha = .1, shape = 20) +
  geom_smooth() +
  facet_grid(rows = vars(metric), cols = vars(rho), scales = "free_y") +
  labs(
    color = "Productivity",
    x = "Species richness",
    y = "Stability component"
  ) +
  scale_x_continuous(
    sec.axis = sec_axis(~ . ,
      name = "Response diversity",
      breaks = NULL,
      labels = NULL)
  )
ggsave(
  filename = here::here("fig", "stability_rich_fw_only.pdf"),
  p_sim_rich_prod,
  scale = 1.8,
  width = 88,
  height = 88 * .6,
  units = "mm"
)
```

```{r}
p_sim_rich_prod_p_only <- sim_rich_prod %>%
  filter(ct_alive == 0) %>%
  mutate(stab_pop = 1/ avg_cv_sp) %>%
  pivot_longer(
    c(stab_com, stab_pop, sync),
    names_to = "metric",
    values_to = "stab") %>%
  mutate(
    metric = var_replacement()[metric],
    rho = rho_replacement()[as.character(rho)]
  ) %>%
  ggplot(aes(x = richness, y = stab, color = factor(productivity))) +
  geom_point(alpha = .1, shape = 20) +
  geom_smooth() +
  facet_grid(rows = vars(metric), cols = vars(rho), scales = "free_y") +
  labs(
    color = "Productivity",
    x = "Species richness",
    y = "Stability component"
  ) +
  scale_x_continuous(
    sec.axis = sec_axis(~ .,
      name = "Response diversity",
      breaks = NULL,
      labels = NULL)
  )
ggsave(
  filename = here::here("fig", "stability_rich_p_only.pdf"),
  p_sim_rich_prod_p_only,
  scale = 1.8,
  width = 88,
  height = 88 * .6,
  units = "mm"
)
```

@@

@@

```{r}
ctrl <- trainControl(method = "cv",  number = 10)

# CV bagged model
bagged_cv <- train(
  stab_com ~ richness + ct_alive + rho + productivity + max_tlvl_alive +
    avg_max_int_alive,
  data = simZ,
  cp = .2,
  maxdepth = 6,
  method = "treebag",
  trControl = ctrl,
  importance = TRUE
  )
plot(varImp(bagged_cv), 20)
plot(bagged_cv)
```

```{r}
m1 <- rpart(
  formula = stab_com ~ richness + ct_alive + rho + productivity + max_tlvl_alive +
    avg_max_int_alive,
  data    = simZ,
  method  = "anova"
  )
rpart.plot(m1)
```
